---
title: "GTFS unification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GTFS unification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(SiteSelection)
```


# Introduction 

Public transit analysis takes advantage of the standardized GTFS format. However, its provision by operator makes it difficult for network aggregated analysis, considering connectivity and multimodality. `SiteSelection::gtfs_unify` proposes a simple solution to this problem, generating an aggregated GTFS file given several instances of these.

# Download GTFS files 
Analyzing public transit frequency is important to understand its territorial coverage and dynamics, both on its spatial and temporal dimensions. SiteSelection provides a simple way to perform this analysis, through SiteSelection::gtfs_frequency() method, explored in this document.


## Data source

The gathering of the GTFS files can be simplified by using public archives like [transitfeeds.com](https://transitfeeds.com/) or [transit.land](https://www.transit.land/).

```{r message=TRUE, warning=FALSE}
data <- read.csv(system.file("extdata", "gtfs_sources_pt.csv", package = "SiteSelection"))
```


This library offers a small database with a compilation of GTFS files for Portuguese operators. It is a CSV file, available at `extdata/gtfs_sources_pt.csv`, and has the following attributes:

- `ID`, a string unique identifier of the region/city the GTFS file applies to;
- `LastUpdate`, the date at which this database entry was last updated;
- `ReferenceDate`, a representative Wednesday that falls within the GTFS calendar;
- `URL`, the URL at which the GTFS file is available;
- `GTFSDocs`, the URL to the page that documents the operator GTFS.


```{r message=TRUE, warning=FALSE}
data <- read.csv(system.file("extdata", "gtfs_sources_pt.csv", package = "SiteSelection"))
summary(data)
data[1:4,]
```

## Download and validate integrity

`SiteSelection` library offers a method to download and fix any integrity violations at GTFS files: `SiteSelection::gtfs_download()`. It fixes inconsistencies on the `stop_times.txt` and missing `shapes.txt` files.

```{r message=TRUE, warning=TRUE}
# DOWNLOAD GTFS and store it locally
gtfs_local <- lapply(data[3:4,]$ID, function(ID) {
  return(gtfs_download(data$URL[data$ID == ID], sprintf("database/transit/%s_gtfs.zip", ID)))
})
```


# Unify GTFS

The unification is performed through `SiteSelection::gtfs_unify()`, producing a single GTFS instance, saved as a ZIP file. Option `generateTransfers` enables the generation of `transfers.txt`, aggregating close stops, even if from different GTFS. 

```{r}
# Load GTFS
library(GTFSwizard)
gtfs_list <- lapply(gtfs_local, function(location) {
  return(GTFSwizard::read_gtfs(location))
})

# Perform unification
gtfs_unify(gtfs_list, "database/transit/gtfs_unified.zip", generateTransfers=TRUE)
```


